---
title: "Prices dataset"
subtitle: "Hamish Patten and Ignacio Acosta"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    css: styles.css
---

<style type="text/css">
h1.title {
  font-size: 50px;
  color: White;
}
h3.subtitle {
  font-size: 20px;
  color: White;
}
<!-- .table-hover > tbody > tr:hover { -->
<!--   background-color: #f7f7e6; -->
<!-- } -->
.main-container {
  width: 95%;
    <!-- max-width: unset; -->
  }
</style>

<div style= "position:relative">
<!-- <div style= "float:left; position:relative; top:-100px;margin-bottom:-100px;margin-left:-10px;"> -->

```{r setup, include=FALSE}
# Setup the local environment
source("./RCode/Data_Wrangling/GetFAOSTAT.R")

  library(dplyr)
  library(magrittr)
  library(tidyverse)
  library(readxl)
  library(ggplot2)
  library(countrycode)
  library(stringr)
  library(pracma)
  library(data.table)
  library(FAOSTAT)

# Some extras for nice and neat tables
knitr::opts_chunk$set(echo = TRUE)
options(kableExtra.latex.load_packages = FALSE,
        knitr.kable.NA = '')


FAOSTAT <- GetFAOSTAT_All(syear = 1990,fyear = 2024)
Production <- FAOSTAT$Prod
Prices <- FAOSTAT$price
```

</div>

## Prices dataset

The **FAOSTAT Producer Prices dataset** gather information related to prices received by farmers for primary crops, live animals and livestock primary products as collected at the point of initial sale (prices paid at the farm-gate). 

- Annual data are provided from 1991.

The variable of interest for our report is **USD/Tonne** which is then converted to **2017 USD/Tonne**.

The dataset contains price observation for the Country-Year-Item combination.

Example: Producer price (measured in 2017 USD/Tonne) of strawberries in Uruguay for the year 2014.

## Work until now

In previous report, the work flow was the following:

1) Production losses were estimated, therefore, we got a dataset containing the variables Country-Year-Item-Production_loss.

2) We left-join the prices dataset.

Here, some **NA's** were created.

### How part of NA's were handled:

Subregional, regional and world **MEAN** prices were computed at an **ITEM - Year** level.

Ex: The imputation dataset contains the mean price of strawberries at the subregional, regional, and world levels for each year.

Then, if a country had an NA in price, it was imputed using the smallest available level of dissagregation.

# Prices dataset descriptives

Items which are present in Production dataset but not in Prices dataset:

```{r, message=FALSE,warning=FALSE, echo=FALSE}
Production%>%
  group_by(
    Item
  )%>%
  summarise(
      n = n(),
      Production = sum(Production,na.rm=TRUE)
      )%>%
  anti_join(
    Prices%>%
      group_by(
        Item
      )%>%
      count(
      )
      ,
      by = "Item"
  )%>%
  arrange(desc(Production))%>%
kableExtra::kable(
  col.names = c(
    "Item",
    "n",
    "Production"
    )
  )
```
There are 79 Item which will have NA after imputation.
